On Extending a Full-Sharing Multithreaded Tabling Design with Batched Scheduling

### 2015-04-17 paper.tex
final version

\begin{abstract}
Tabling is an implementation technique that overcomes some limitations
of traditional Prolog systems in dealing with redundant
sub-computations and recursion. Multithreading in Prolog is the
ability to concurrently perform computations, in which each
computation runs independently but shares the program clauses. When
multithreading is combined with tabling, we have the best of both
worlds, since we can exploit the combination of higher procedural
control with higher declarative semantics.  To support this
combination, the Yap system has at engine level, multiple designs that
vary from a No-Sharing design, where each thread allocates fully
private tables for each new tabled subgoal called during its
computation, to a Full-Sharing design, where threads share the
complete table space. Arguably, batched scheduling is one of the most
successful tabling scheduling strategies. In this work, we propose an
extension of the Full-Sharing design to support batched
scheduling. Experimental results show that, despite the extra data
structures required to support the extension, the FS design remains
quite competitive.

@@

With the FS design, all tables are shared. Thus, since several threads
can be inserting answers in the same table, when an answer already
exists, it is not possible to determine if the answer is new or
repeated for a certain thread without further support. However, for
local scheduling this is not a problem because, for repeated and new
answers, local scheduling always fails. The problem is with batched
scheduling that requires further support to work with the FS design
since, with batched scheduling, only repeated answers should fail. We
have thus to detect when a new found answer is also a new answer for
the thread at hand in order to ensure that new answers for a given
subgoal call are correctly propagated one and only once to all
thread's calls.

@@

In Yap, a thread views its tables as private but, at the engine level,
it uses a common table space where tables are shared among all
threads. Yap implements three designs for concurrent tabling support
that vary from a \emph{No-Sharing} (NS) design, where each thread
allocates fully private tables for each new tabled subgoal called
during its computation, to a \emph{Full-Sharing} (FS) design, where
threads share the complete table space. Figure~\ref{fig_table_space}
shows Yap's multithreaded table space organization for the NS and FS
designs, where an interface layer abstracts the design being used at
the engine level. The figure illustrates the main differences between
the two designs for a situation where several threads are evaluating
the same tabled subgoal call $call\_i$.

@@

Local scheduling schedules the evaluation of a program in a
breath-first manner. It favors backtracking first with completion
instead of forward execution, leaving the consumption of answers for
last. Local scheduling only allows a \emph{Cluster of Dependent
  Subgoals} (CDS) to return answers after a fixpoint point has been
reached~\cite{Freire-96}.

In other words, local scheduling tries
to keep CDS as minimal as possible. When new answers are found, they
are added to the table space and the computation fails as consequence,
tabled subgoals inside a CDS propagate their answers to outside the
CDS only after its completion point is found.

Local scheduling causes a sooner completion of subgoals, which creates
less complex dependencies between them.

On the other hand, batched scheduling schedules the evaluation of a
program in a depth-first manner. It favors the forward execution first
instead of backtracking, leaving the consumption of answers and
completion for last. It thus tries to delay the need to move around
the search tree by batching the return of answers. When new answers
are found for a particular tabled subgoal, they are added to the table
space and the execution continues. For some situations, this results
in creating dependencies to older subgoals, therefore enlarging the
current CDS~\cite{Sagonas-98} and delaying the completion point to an
older generator node.

@@

In this section, we show how we have extended the FS design to support
batched scheduling. The usage of FS design with batched scheduling,
requires further support than with local scheduling, since when a new
answer is found it must be immediately propagated, and we have to
ensure that the propagation occurs on all subgoal calls once and only
once. To do so, next we show a procedure that, for every subgoal call
of every thread, keeps track of all the answers that were already
propagated.

@@

The key idea of our approach, which we named Privately-consumed Answer
Chaining (PAC), is then to extend the \emph{FS} design with batched
scheduling, by chaining privately for each subgoal call, the answers
that were already consumed by a thread. Since the procedure is
private, it will only affect the thread that is doing it. At the end,
when the evaluation is complete, i.e, when a subgoal call is marked as
complete, we put one of the private chain as public, so that from that
point on all threads can use that chain in complete (only reading)
mode.

Figure~\ref{fig_tabtries_pcc} shows the key data structures for
supporting the implementation of the PAC procedure during the
evaluation of a tabled subgoal call $i$ using the FS design. %% The FS
%% design, uses a subgoal entry data structure to store common
%% information for a subgoal call and a subgoal frame (\emph{SF}) data
%% structure to store private information about the execution of each
%% thread.
The PAC procedure works at the subgoal frame level, which is
private to each thread.

\begin{figure}[!ht]
%\vspace{-\intextsep}
\centering
\includegraphics[width=10.5cm]{figures/pcc.pdf}
\caption{The PAC procedure using (a) private chaining and (b) public
  chaining}
\label{fig_tabtries_pcc}
%\vspace{-1.0\intextsep}
\end{figure}

Figure~\ref{fig_tabtries_pcc}(a) shows then a situation where two
threads, $T_1$ and $T_{k-1}$, are sharing the same subgoal entry call
$i$ when the subgoal is still under evaluation, i.e., the subgoal is
not yet complete. The current state of the evaluation shows an answer
trie with $3$ answers found for the subgoal call $i$. For the sake of
simplicity, we are omitting the internal answer trie nodes and we are
only showing the leaf nodes nodes $LN_1$, $LN_2$ and $LN_3$, in the
figure. %% With the PAC procedure, the leaf nodes are not chained in the
%% answer trie data structure. Now, the chaining process is done
%% privately, and for that, we use the subgoal frame structure of each
%% thread.
On the subgoal frame structure we added a new field, called
\emph{consumer}, to store the answers found within the execution of
the thread. In order to minimize the impact of the \emph{PAC}
optimization, each node within the new consumer answer structure has
two fields: (i) an entry pointer, which points to the corresponding
leaf node in the answer trie data structure; (ii) a next pointer to
chain the answers within the consumer structure. To maintain a good
performance, when the number of nodes exceeds a certain threshold, we
use a hash trie mechanism design similar to the one presented in the
work~\cite{Areias-ijpp15}. However, since this mechanism is private to
each thread, it does not require any of the tools that were necessary
to support concurrency, thus we have removed them from the mechanism.

%% In particular, on each hash trie level, we
%% have removed the tools necessary to support concurrency, such as
%% useless pointers and compare-and-swap operations. We have chosen this
%% hashing mechanism, because it showed a good balance between lookup and
%% insert operations~\cite{Areias-ijpp15}, but the major reason was
%% mostly because of the integration in the TabMalloc memory
%% allocator~\cite{Areias-12b}.

Going back to Figure~\ref{fig_tabtries_pcc}(a), the consumer answer
structures represent then two different situations where threads can
be evaluating a subgoal call. Thread $T_1$ has only found one answer
and it is using a direct consumer answer chaining to access the node
$LN_1$. Thread $T_{k-1}$ was already found three answers for the
subgoal call and it is already using the hash trie mechanism within
its consumer answer structure. The consumer nodes are chained between
themselves, thus that consumer nodes belonging to thread $T_{k-1}$ can
consume the answers as in the original mechanism.

Figure~\ref{fig_tabtries_pcc}(b) shows the state of the subgoal call
after completion. When a thread $T$ completes a subgoal call, it frees
its private consumer structures, but before doing that, it checks
whether another thread as already marked the subgoal as completed. If
no other thread has done that, then thread $T$ not only follows its
private chaining mechanism as it would for freeing its private nodes,
but also, follows the pointers to the answer trie leaf nodes in order
to create a chain inside the answer trie. Since this procedure is done
inside a critical region, no more than one thread can be doing this
chaining process. Thus, in Figure~\ref{fig_tabtries_pcc}(b), we are
showing a situation where the subgoal call is completed and both
threads $T_1$ and $T_{k-1}$ have already removed their consumer answer
structures and chained the leaf nodes inside the answer trie.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Implementations Details}

At the tabling engine, the major difference between local and batched
scheduling is in the tabling operation \emph{tabled new answer}, where
we decide what to do when an answer is found during the
evaluation. This operation checks whether a newly found answer is
already in the corresponding answer trie structure and, if not,
inserts it. Algorithm~\ref{alg_table_new_answer_batched} shows how we
have extended this operation to support FS design with batched
scheduling.

@@

The algorithm receives two arguments: the new answer found during the
evaluation (\emph{ANS}) and the subgoal frame which corresponds to the
call at hand (\emph{SF}). The algorithm begins by checking/inserting
the given \emph{ANS} into the answer trie structure, which will return
the leaf node for the path representing \emph{ANS} (line 1).

Then, it checks/inserts the given \emph{leaf} node into the private
consumer chain for the current thread, which will return the
corresponding chain node (line 2). Next in line 3, it tests whether
the chain node already existed in the consumer chain, i.e., if it was
inserted or not by the current check/insert operation in order to
return failure (line 4), or it proceeds with marking the answer
\emph{ANS} has found (line 6). At the end (lines 7 to 10), it returns
failure if local scheduling is active (line 8), otherwise, the batched
scheduling is active, thus it propagates the answer \emph{ANS} (line
10).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Performance Analysis}

We now present experimental results about the usage of the batched
scheduling on FS design. To put results in perspective, we will be
showing also the results for the NS design. The environment for our
experiments was a machine with 32-Core AMD Opteron (TM) Processor 6274
(2 sockets with 16 cores each) with 32G of main memory, running the
Linux kernel is the 3.16.7-200.fc20.x86\_64 with Yap 6.3.

For the experimentation, we used our memory allocator described in the
work~\cite{Areias-12b}, later named \emph{TabMalloc}, and we the same
five sets of benchmarks. These benchmarks create \emph{worst case}
scenarios, where we are able to show the lowest bounds in terms of
performance that each design might achieve when applied/used with
other real world applications/programs.

@@

\setlength{\tabcolsep}{12pt}
\begin{table}[!ht]
%\vspace{-\intextsep}
\centering
%% \caption{Overhead ratios, when compared with the NS design with $1$
%%   thread (running with local scheduling without TabMalloc) for the NS
%%   and FS designs (with TabMalloc), when running 1, 8, 16, 24 and 32
%%   threads with local and batched scheduling on the five sets of
%%   benchmarks (best ratios by row and by design for the Minimum,
%%   Average and Maximum are in bold)}
\caption{Overhead ratios, when compared with the NS design with 1
  thread for the NS and FS designs, when running 1, 8, 16, 24 and 32
  threads with local and batched scheduling (lowest is the best and it
  is marked in bold by row and by design for the Minimum, Average and
  Maximum)}
%\vspace{-0.5\intextsep}
%\scalebox{1.00} {
\begin{tabular}{ll|cc|cc}
\hline\hline
\multicolumn{2}{c|}{\multirow{2}{*}{\bf Threads}} &
\multicolumn{2}{c|}{\multirow{1}{*}{\bf NS}} &
\multicolumn{2}{|c}{\multirow{1}{*}{\bf FS}}\\ \cline{3-6}
& 
& \multicolumn{1}{c}{\bf Local}
& \multicolumn{1}{c}{\bf Batched}
& \multicolumn{1}{|c}{\bf Local}
& \multicolumn{1}{c}{\bf Batched}\\
\hline
\multirow{4}{*}{\bf 1}
& {\bf Min }& {\bf 0.53}& 0.55& 1.01& {\bf 0.95}\\
& {\bf Avg }& {\bf 0.78}& 0.82& {\bf 1.30}& 1.46\\
& {\bf Max }& 1.06& {\bf 1.05}& {\bf 1.76}& 2.33\\
& {\bf StD }& 0.15& 0.14& 0.22& 0.44\\
\hline
\multirow{4}{*}{\bf 8}
& {\bf Min }& 0.66& {\bf 0.63}& 1.16&{\bf  0.99}\\
& {\bf Avg }& {\bf 0.85}& 0.88& {\bf 1.88}& 1.95\\
& {\bf Max }& {\bf 1.12}& 1.14& {\bf 2.82}& 3.49\\
& {\bf StD }& 0.13& 0.14& 0.60& 0.79\\
\hline
\multirow{4}{*}{\bf 16}
& {\bf Min }& 0.85& {\bf 0.75}& 1.17& {\bf 1.06}\\
& {\bf Avg }& {\bf 0.98}& 1.00& {\bf 1.97}& 2.08\\
& {\bf Max }& {\bf 1.16}& 1.31& {\bf 3.14}& 3.69\\
& {\bf StD }& 0.09& 0.17& 0.65& 0.83\\
\hline
\multirow{4}{*}{\bf 24}
& {\bf Min }& {\bf 0.91}& 0.93& 1.16& {\bf 1.09}\\
& {\bf Avg }& {\bf 1.15}& 1.16& {\bf 2.06}& 2.19\\
& {\bf Max }& 1.72& {\bf 1.60}& {\bf 3.49}& 4.08\\
& {\bf StD }& 0.20& 0.21& 0.70& 0.91\\
\hline
\multirow{4}{*}{\bf 32}
& {\bf Min }& 1.05& {\bf 1.04}& 1.33& {\bf 1.26}\\
& {\bf Avg }& 1.51& {\bf 1.49}& {\bf 2.24}& 2.41\\
& {\bf Max }& {\bf 2.52}& 2.63& {\bf 3.71}& 4.51\\
& {\bf StD }& 0.45& 0.45& 0.74& 1.02\\
\hline\hline
\end{tabular}%}
\label{tab_batched_overhead}
%\vspace{-0.7\intextsep}
\end{table}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions and Further Work}
We have presented a simple and novel approach for combining the FS
design with batched scheduling. Experimental results showed that on
average, the extra structures required for the combination do not seem
to have a big impact in the performance. Further work will include
experimentation in other real world applications.

@@

%\section*{Acknowledgments}

>>>>

\begin{abstract}
Tabling is an implementation technique that overcomes some limitations
of traditional Prolog systems in dealing with redundant
sub-computations and recursion. Multithreading in Prolog is the
ability to concurrently perform computations, in which each
computation runs independently but shares the program clauses. When
multithreading is combined with tabling, we have the best of both
worlds, since we can exploit the combination of higher procedural
control with higher declarative semantics. To support this
combination, the Yap Prolog system has, at engine level, multiple
designs that vary from a No-Sharing design, where each thread
allocates fully private tables for each new tabled subgoal called
during its computation, to a Full-Sharing design, where threads share
the complete table space. Arguably, batched scheduling is one of the
most successful tabling scheduling strategies. In this work, we
propose an extension of the Full-Sharing design to support batched
scheduling. Experimental results show that, despite the extra data
structures required to support the extension, the FS design remains
quite competitive.\\

@@

With the FS design, all tables are shared. Thus, since several threads
can be inserting answers in the same table, when an answer already
exists, it is not possible to determine if the answer is new or
repeated for a certain thread without further support. However, for
local scheduling this is not a problem because, for repeated and new
answers, local scheduling always fails. The problem is with batched
scheduling that requires further support to work with the FS design
since, with batched scheduling, only repeated answers should fail. We
have thus to detect when a newly found answer is also a new answer for
the thread at hand in order to ensure that new answers for a given
subgoal call are correctly propagated one and only once to all
thread's repeated calls.

@@

In Yap, a thread views its tables as private but, at the engine level,
it implements three designs for concurrent tabling support that vary
from a \emph{No-Sharing} (NS) design, where each thread allocates
fully private tables for each new tabled subgoal called during its
computation, to a \emph{Full-Sharing} (FS) design, where threads share
the complete table space. Figure~\ref{fig_table_space} shows Yap's
multithreaded table space organization for the NS and FS designs,
where an interface layer abstracts the design being used at the engine
level. The figure illustrates the main differences between the two
designs for a situation where several threads are evaluating the same
tabled subgoal call $call\_i$.

@@

Local scheduling schedules the evaluation of a program in a
breath-first manner. It favors backtracking first with completion
instead of forward execution, leaving the consumption of answers for
last. Local scheduling only allows a \emph{Cluster of Dependent
  Subgoals} (CDS) to return answers after a fix-point has been
reached~\cite{Freire-96}. In other words, local scheduling tries to
keep a CDS as minimal as possible, thus creating less complex
dependencies between subgoals, which causes a sooner completion of
subgoals.

On the other hand, batched scheduling schedules the evaluation of a
program in a depth-first manner. It favors the forward execution first
instead of backtracking, leaving the consumption of answers and
completion for last. It thus tries to delay the need to move around
the search tree by batching the return of answers. When new answers
are found for a particular tabled subgoal, they are added to the table
space and the execution continues. For some situations, this results
in creating dependencies to older subgoals, therefore enlarging the
current CDS and delaying the fix-point that guarantees that all
dependent subgoals in a CDS are completely
evaluated~\cite{Sagonas-98}.

@@

In this section, we describe our proposal to support the combination
of batched scheduling with the FS design. Remember that batched
scheduling requires further support to work with the FS design
because, with batched scheduling, we have to detect when a newly found
answer is also a new answer for the thread at hand in order to ensure
that new answers for a given subgoal call are correctly propagated one
and only once to all thread’s repeated calls. To solve that, we propose
an extension to the table space data structures, which we named
\emph{Private Answer Chaining} (PAC), as a way to keep track, per
thread and subgoal call, of the answers that were already found and
propagated to the thread's repeated calls.

@@

The PAC procedure works at the subgoal frame level, which is private
to each thread. The key idea of the PAC data structure is then to
extend subgoal frames with an auxiliary private chaining of answers
for each subgoal call, in order to keep track of the answers already
found for the call. Later, if a thread completes a subgoal's
evaluation, i.e, if the subgoal's table is marked as complete, its PAC
is made public, so that from that point on all threads can use that
chain in complete (only reading) mode. Figure~\ref{fig_tabtries_pcc}
illustrates the new data structures involved in the implementation of
our PAC's proposal for a situation where several threads are
evaluating the same tabled subgoal call $call\_i$.

\begin{figure}[!ht]
%\vspace{-\intextsep}
\centering
\includegraphics[width=10.5cm]{figures/pcc.pdf}
\caption{PAC's data structures for (a) private and (b) public
  chaining}
\label{fig_tabtries_pcc}
%\vspace{-1.0\intextsep}
\end{figure}

Figure~\ref{fig_tabtries_pcc}(a) shows then a situation where two
threads, $T_1$ and $T_{k-1}$, are sharing the same subgoal entry for
the call $call\_i$ when the subgoal is still under evaluation, i.e.,
the subgoal is not yet complete. The current state of the evaluation
shows an answer trie with 3 answers found for the subgoal call
$call\_i$. For the sake of simplicity, we are omitting the internal
answer trie nodes and we are only showing the leaf nodes nodes $LN_1$,
$LN_2$ and $LN_3$, in the figure.

With PAC support, the leaf nodes are not chained in the answer trie
data structure, as usual. Now, the chaining process is done privately,
and for that, we use the subgoal frame structure of each thread. On
the subgoal frame structure we added a new field, called
\emph{Answers}, to store the answers found within the execution of the
thread. In order to minimize PAC's impact, each answer node in the
private chaining has only two fields: (i) an entry pointer, which
points to the corresponding leaf node in the answer trie data
structure; and (ii) a next pointer to chain the nodes in the private
chaining. To maintain a good performance, when the number of answer
nodes exceeds a certain threshold, we use a hash trie mechanism design
similar to the one presented in the
work~\cite{Areias-ijpp15}. However, since this mechanism is private to
each thread, it does not require any concurrency support.

Going back to Fig.~\ref{fig_tabtries_pcc}(a), the PAC's data
structures represent then two different situations where threads can
be evaluating a subgoal call. Thread $T_1$ has only found one answer
and it is using a direct answer chaining to access the node
$LN_1$. Thread $T_{k-1}$ was already found three answers for the
subgoal call and it is using the hash trie mechanism within its
private chaining. Please note that the answer nodes are still chained
between themselves, thus that repeated calls belonging to thread
$T_{k-1}$ can consume the answers as in the original mechanism.

Figure~\ref{fig_tabtries_pcc}(b) shows the state of the subgoal call
after completion. When a thread $T$ completes a subgoal call, it frees
its private consumer structures, but before doing that, it checks
whether another thread as already marked the subgoal as completed. If
no other thread has done that, then thread $T$ not only follows its
private chaining mechanism as it would for freeing its private nodes,
but also, follows the pointers to the answer trie leaf nodes in order
to create a chain inside the answer trie. Since this procedure is done
inside a critical region, no more than one thread can be doing this
chaining process. Thus, in Fig.~\ref{fig_tabtries_pcc}(b), we are
showing the situation where the subgoal call is completed and both
threads $T_1$ and $T_{k-1}$ have already chained the leaf nodes inside
the answer trie and removed their private chaining structures.

\subsection{Implementations Details}

At the tabling engine, the major difference between local and batched
scheduling is in the tabling operation \emph{tabled new answer}, where
we decide what to do when an answer is found during the
evaluation. This operation checks whether a newly found answer is
already in the corresponding answer trie structure and, if not,
inserts it. Algorithm~\ref{alg_table_new_answer_batched} shows how we
have extended this operation to support the FS design with batched
scheduling.

@@

The algorithm receives two arguments: the newly answer found during
the evaluation (\emph{ANS}) and the subgoal frame which corresponds to
the call at hand (\emph{SF}). The algorithm begins by
checking/inserting the given \emph{ANS} into the answer trie
structure, which will return the leaf node for the path representing
\emph{ANS} (line 1). Then, it checks/inserts the given \emph{leaf}
node into the private chaining for the current thread, which will
return the corresponding answer chain node (line 2). Next in line 3,
it tests whether the answer chain node already existed in the chain,
i.e., if it was inserted or not by the current check/insert operation
in order to return failure (line 4), or it proceeds with marking the
answer \emph{ANS} has found (line 6). At the end (lines 7 to 10), it
returns failure if local scheduling is active (line 8), otherwise,
batched scheduling is active, and it proceeds by propagating the
answer \emph{ANS} to the current execution environment (line 10).

\section{Performance Analysis}

We now present experimental results about the usage of the FS design
with batched scheduling. To put results in perspective, we will be
showing also the results for local scheduling and for the NS
design.

The environment for our experiments was a machine with 32-Core AMD
Opteron (TM) Processor 6274 (2 sockets with 16 cores each) with 32G of
main memory, running the Linux kernel is the 3.16.7-200.fc20.x86\_64
with Yap Prolog 6.3. For the experiments, we used the \emph{TabMalloc}
memory allocator~\cite{Areias-12b} and the set of benchmarks described
in~\cite{Areias-12b}. These benchmarks create \emph{worst case
  scenarios}, where we are able to show the lowest bounds in terms of
performance that each design might achieve when applied/used with
other real world applications/programs.

@@

\setlength{\tabcolsep}{12pt}
\begin{table}[!ht]
\centering
\caption{Overhead ratios, when compared with the NS design with $1$
  thread (running with local scheduling and without TabMalloc) for the NS
  and FS designs (with TabMalloc), when running 1, 8, 16, 24 and 32
  threads with local and batched scheduling (best ratios by row and by
  design for the Minimum, Average and Maximum are in bold)}
\begin{tabular}{ll||cc|cc}
\multicolumn{2}{c||}{\multirow{2}{*}{\bf Threads}} &
\multicolumn{2}{c|}{\multirow{1}{*}{\bf NS}} &
\multicolumn{2}{|c}{\multirow{1}{*}{\bf FS}}\\ \cline{3-6}
& 
& \multicolumn{1}{c}{\bf Local}
& \multicolumn{1}{c}{\bf Batched}
& \multicolumn{1}{|c}{\bf Local}
& \multicolumn{1}{c}{\bf Batched}\\
\hline\hline
\multirow{4}{*}{\bf 1}
& {\bf Min }& {\bf 0.53}& 0.55& 1.01& {\bf 0.95}\\
& {\bf Avg }& {\bf 0.78}& 0.82& {\bf 1.30}& 1.46\\
& {\bf Max }& 1.06& {\bf 1.05}& {\bf 1.76}& 2.33\\
& {\bf StD }& 0.15& 0.14& 0.22& 0.44\\
\hline
\multirow{4}{*}{\bf 8}
& {\bf Min }& 0.66& {\bf 0.63}& 1.16&{\bf  0.99}\\
& {\bf Avg }& {\bf 0.85}& 0.88& {\bf 1.88}& 1.95\\
& {\bf Max }& {\bf 1.12}& 1.14& {\bf 2.82}& 3.49\\
& {\bf StD }& 0.13& 0.14& 0.60& 0.79\\
\hline
\multirow{4}{*}{\bf 16}
& {\bf Min }& 0.85& {\bf 0.75}& 1.17& {\bf 1.06}\\
& {\bf Avg }& {\bf 0.98}& 1.00& {\bf 1.97}& 2.08\\
& {\bf Max }& {\bf 1.16}& 1.31& {\bf 3.14}& 3.69\\
& {\bf StD }& 0.09& 0.17& 0.65& 0.83\\
\hline
\multirow{4}{*}{\bf 24}
& {\bf Min }& {\bf 0.91}& 0.93& 1.16& {\bf 1.09}\\
& {\bf Avg }& {\bf 1.15}& 1.16& {\bf 2.06}& 2.19\\
& {\bf Max }& 1.72& {\bf 1.60}& {\bf 3.49}& 4.08\\
& {\bf StD }& 0.20& 0.21& 0.70& 0.91\\
\hline
\multirow{4}{*}{\bf 32}
& {\bf Min }& 1.05& {\bf 1.04}& 1.33& {\bf 1.26}\\
& {\bf Avg }& 1.51& {\bf 1.49}& {\bf 2.24}& 2.41\\
& {\bf Max }& {\bf 2.52}& 2.63& {\bf 3.71}& 4.51\\
& {\bf StD }& 0.45& 0.45& 0.74& 1.02\\
\end{tabular}
\label{tab_batched_overhead}
\end{table}

@@

@@

\section{Conclusions and Further Work}

We have presented a simple and novel approach for combining the FS
design with batched scheduling. Experimental results showed that on
average, the extra structures required for the combination do not seem
to have a big impact in the performance. Further work will include
experimentation in other real world applications.

\section*{Acknowledgments}

This work is partially funded by the ERDF (European Regional
Development Fund) through the COMPETE Programme and by FCT (Portuguese
Foundation for Science and Technology) within projects SIBILA
(NORTE-07-0124-FEDER-000059) and PEst
(FCOMP-01-0124-FEDER-037281). Miguel Areias is funded by the FCT grant
SFRH/BD/69673/2010.
